---
title: "Data Challenge"
author: "Jennifer Dusendang"
date: "12/14/2019"
output: html_document


Prior to 1990 did docs who passed their on their intitial try have better evaluation scores?

```{r a}
#read in data 
library(questionr)
library(ggplot2)
docinfo <- read.csv("C:/Users/jduse_000/Documents/xx/match_data.csv") 
docscores <- read.csv("C:/Users/jduse_000/Documents/xx/match_data_year_data.csv")

#merge datasets 
fulldata <- merge(docinfo, docscores, by.x="physician_uid")

#define pass / fail groups 
#if their residency year = certification year then we assume initial passing 
#first only consider those with certification ending before 1989 (need one year of data)
before1990 <- fulldata[which(fulldata$board_cert_year < 1989),]

before1990$initial_exam <- ifelse(before1990$training_year == before1990$board_cert_year, "PASSED", "FAILED")

sum(before1990$initial_exam == 'PASSED')
sum(before1990$initial_exam == 'FAILED')
before1990$guideline_risk <- rowMeans(before1990[, c(4:35)], na.rm = TRUE)

#ttest 
t.test(before1990$guideline_risk~before1990$initial_exam)

```

Look at initial evaluation score to see if docs after the change had higher initial evaluation scores 

```{r b}

#changes to certification process 

#initial evaluation score 
for (row in 1:nrow(fulldata)){
  mylist <- as.numeric(fulldata[row, 4:35])
  myfirst <- gdata::first(na.omit(mylist))
  fulldata[row,36] <- myfirst
}
fulldata <- questionr::rename.variable(fulldata, "V36", "initial_score")
fulldata$prepost <- ifelse(fulldata$board_cert_year < 1990, "PRE", "POST")
t.test(fulldata$initial_score~fulldata$prepost)


```

Did recertification help in increasing scores?

```{r c}

#av performance score by year for graph 
fulldata$guideline_risk <- rowMeans(fulldata[, c(4:35)], na.rm = TRUE)
fulldata$prepost <- ifelse(fulldata$board_cert_year < 1990, "PRE", "POST")
t.test(fulldata$guideline_risk~fulldata$prepost)
min(fulldata[which(fulldata$prepost == "PRE"),]$guideline_risk)
min(fulldata[which(fulldata$prepost == "POST"),]$guideline_risk)

ggplot(data = fulldata, aes(x = factor(board_cert_year), y = guideline_risk, ymin = 0, ymax = 100)) + stat_summary(fun.y="mean", geom="point", size = 3) + xlab("Board Certification Year") + ylab("Average of Evaluation Scores Over Physician's Career") + theme_classic() 

#av performance score by those who failed recertification 
after1990 <- fulldata[which(fulldata$board_cert_year >= 1990),]
after1990$activeyears <- rowSums(!is.na(after1990[, c(4:35)]))
after1990$contcertified <- ifelse(after1990$activeyears == 2002 - after1990$board_cert_year, "CONT", "NON-CONT")
t.test(after1990$guideline_risk~after1990$contcertified)

```

Did more people fail initial certification after the change?


```{r d}

#have more people failed the exam after 1990?
fulldata$initial_exam <- ifelse(fulldata$training_year == fulldata$board_cert_year, "PASSED", "FAILED")
table(fulldata$prepost, fulldata$initial_exam)

#overall count of certified physicans by year 
table(fulldata$board_cert_year)

#overal count of number of evaluations performed by year 
colSums(!is.na(fulldata[, c(4:35)]))
```